---
title: "bridges_homework"
author: "Yifan Zhang"
date: "2021/2/4"
output: github_document
---

## Prerequisite

```{r}
#install.packages("tidyverse")
#install.packages("backports")
library(tidyverse)
```

## Import data


```{r}
AllBridgeData <- read.csv("bridge_data.txt")
```

## Pick some variables

Make a file with bridge ID, year, fips codes, condition ratings, and a few other variables. According to the data description, we need to find item 8 (structure Number), item 27 (year), item 1 (state code), item 3(county code), item 4 (place code), and item 58 - 62 (condition ratings). I am also interested in item 2 (highway agency district), item 16 (latitude), and item 17 (longitude).

```{r}
BridgeData <- select(AllBridgeData, STRUCTURE_NUMBER_008, YEAR_BUILT_027, DECK_COND_058, SUPERSTRUCTURE_COND_059, SUBSTRUCTURE_COND_060, CHANNEL_COND_061, CULVERT_COND_062, STATE_CODE_001, COUNTY_CODE_003, PLACE_CODE_004, HIGHWAY_DISTRICT_002, LAT_016, LONG_017)
head(BridgeData)
write.csv(BridgeData, "bridge_data_new", row.names = FALSE)
#show the file I made
NewData <- read.csv("bridge_data_new")
head(NewData)
```

## Make a plot

In the warm-up section, we are required to make a plot. I decide to make a plot with latitude and longitude.

```{r}
plot(BridgeData$LAT_016, BridgeData$LONG_017, xlab = "latitude", ylab = "longitude", main = "Latitude&Longitude Map")
```

This is a relatively easy visualization. We expect to see where most of these bridges locate.

## Project Idea

## Subset of "interesting" bridges

I intend to make a subset of all Wisconsin bridges. According to the data description, the FIPS code for Wisconsin State (Item 1) is 55.

```{r}
WisconsinBridges <- BridgeData[which(BridgeData$STATE_CODE_001 == 55), ]
head(WisconsinBridges)
```

## Analyze and visualization

Track how those bridges have changed over time. Make an interactive visualization that helps to tell a story. I intend to build x-axis with condition ratings (58-62) and y-axis with build year (item 27).

```{r}
library(ggplot2)
#install.packages("ggpubr")
library(ggpubr)

a <- ggplot(WisconsinBridges, aes(DECK_COND_058, YEAR_BUILT_027)) + 
  labs(title = "Year vs Deck Condition", x = "Deck Condition", y = "Year Built") +
  ylim(1850, 2020) +
  geom_point(colour = "darkblue")

b <- ggplot(WisconsinBridges, aes(SUPERSTRUCTURE_COND_059, YEAR_BUILT_027)) + 
  labs(title = "Year vs Superstructure Condition", x = "SuperStructure Condition", y = "Year Built") +
  ylim(1850, 2020) +
  geom_point(colour = "gold")

c <- ggplot(WisconsinBridges, aes(SUBSTRUCTURE_COND_060, YEAR_BUILT_027)) + 
  labs(title = "Year vs Substructure Condition", x = "Substructure Condition", y = "Year Built") +
  ylim(1850, 2020) +
  geom_point(colour = "darkcyan")

d <- ggplot(WisconsinBridges, aes(CHANNEL_COND_061, YEAR_BUILT_027)) + 
  labs(title = "Year vs Channel Condition", x = "Channel and Channel Protection", y = "Year Built") +
  ylim(1850, 2020) +
  geom_point(colour = "firebrick")

e <- ggplot(WisconsinBridges, aes(CULVERT_COND_062, YEAR_BUILT_027)) + 
  labs(title = "Year vs Culverts Condition", x = "Culverts Condition", y = "Year Built") +
  ylim(1850, 2020) +
  geom_point(colour = "seagreen")

ggarrange(a, b, c, d, e, ncol = 3, nrow = 2)
```

## Conclusion

From the above visualization, we found that:

1.The ggplot of deck condition is similar to the ggplot of superstructure condition. These two variables may be related to each other.

2.Overall the channel condition is the best, because the minimum of channel condition of bridges in WI is 4.
